### 2011-04-26 draft2.tex
Simplifying once again...

\usepackage{stmaryrd}
\usepackage{fullpage}

@@

\newcommand{\unr}{\texttt{UNR}}
\newcommand{\bad}{\texttt{BAD}}
\newcommand{\any}{\texttt{Any}}
\newcommand{\ok}{\texttt{Ok}}
\newcommand{\mand}{\mbox{ and }}
\newcommand{\motherwise}{\mbox{ otherwise.}}

\begin{eqnarray*}
  x,D &::=& String & (Variables, Data)\\
  e &::=& x \mid (f~\overline{e}) \mid (D~\overline{e}) & (Expressions) \\
  &&\mid \mbox{case $e$ of } [(p_1 \to e_1) \dots (p_n \to e_n)]   & \\ 
  &&\mid \bad \mid \unr & \\
  p &::=& \mbox{let } f~\overline{x} = e ::: t,p \mid \epsilon & (Programs) \\
  \Gamma &::=& (f \mapsto t),\Gamma \mid (f~\vec{x} \mapsto e), \Gamma \mid \epsilon & (Contexts) \\
  t &::=& \{x \mid e \} \mid x:t \to t \mid \any & (Contracts) \\
  \phi &::=& \top \mid \bot \mid x \mid \lnot \phi \mid \phi \land \phi \mid \forall x, \phi & (FOF) \\
\end{eqnarray*}

@@

$t \geq t' \iff \{ e | x \in t' \} \subset \{ e | x \in t \}$ ($t$ is more restrictive than $t'$)
\section{Encoding}

We define two different translations: one for expressions and one for programs.

\subsection{Expressions}
\begin{itemize}
\item $\llbracket \Gamma \vdash \bad \in t \rrbracket = \bot$
\item $\llbracket \Gamma \vdash \unr \in t \rrbracket = \top$
\item \begin{array}{rcll}
  \llbracket \Gamma \vdash x \in t \rrbracket &=& \top & \mbox{ if } x \mapsto t' \in \Gamma \\
  && \bot & \mbox{ otherwise.} 
  \end{array}
\item \begin{array}{rcll}
  \llbracket \Gamma \vdash f~\overline{(e_1,\dots,e_n)} ::: t_1 \to \dots \to t_n \to t_c \in t_c \rrbracket &=& \bigwedge_{1 \leq i \leq n}{\llbracket \Gamma \vdash e_i \in t_i \rrbracket} \\
  \llbracket \Gamma \vdash f~\overline{(e_1,\dots,e_n)} ::: t \in t' \rrbracket &=& \bot \mbox{ ie type mismatch.}\\
  \end{array}
\item $\llbracket \Gamma \vdash D~\overline{e} \in t \rrbracket = \overline{e} \neq \bad \to \llbracket \Gamma \vdash D~\overline{e} \in \ok \rrbracket \land ...$
\item $\llbracket \Gamma \vdash \mbox{case } e \mbox{ of } [(p_1 \to e_1),\dots,(p_n \to e_n)] \in t \rrbracket = e \neq \bad \land \forall \vec{z}~p_i[\vec{z}] = e \to \llbracket \Gamma, TODO \vdash e_i \in t \rrbracket$
\end{itemize}

$$\llbracket \Gamma \vdash e \in t \rrbracket = \phi \iff (\phi = \top \iff \Gamma,e \mbox{ satisfies } t)$$

@@

\begin{array}{rcl}
\llbracket \Gamma \vdash \mbox{ let } f~\overline{x} = e:::t,p \rrbracket &=& \forall f^\star, \forall_{1 \leq i \leq n} x_i, \llbracket \Gamma, x_1 \mapsto t_1,\dots,x_n \mapsto t_n, f^\star \mapsto t \vdash e[f^\star/f] \in t_c \rrbracket \\
&& \land \llbracket \Gamma, f \mapsto t \vdash p \rrbracket
\end{array}

>>>>

\usepackage{stmaryrd}
\usepackage{amsmath}
\usepackage{fullpage}

@@

\newcommand{\unr}{\texttt{UNR}}
\newcommand{\unrphi}{\texttt{UNR}_\phi}
\newcommand{\bad}{\texttt{BAD}}
\newcommand{\badphi}{\texttt{BAD}_\phi}
\newcommand{\any}{\texttt{Any}}
\newcommand{\ok}{\texttt{Ok}}
\newcommand{\mand}{\mbox{ and }}
\newcommand{\motherwise}{\mbox{ otherwise.}}

\begin{eqnarray*}
  x,D &::=& String & (Variables, Data)\\
  e &::=& x \mid f(\overline{e}) \mid D(\overline{e}) & (Expressions) \\
  &&\mid \bad \mid \unr & \\
  p &::=& \mbox{let $f~\overline{x} = $ case $e$ of } [(p_1 \to e_1) \dots (p_n \to e_n)] ::: t,p   & (Programs) \\ 
  &&\mid \mbox{let } f~\overline{x} = e ::: t,p \mid \epsilon \\
  \Gamma &::=& (f \mapsto t),\Gamma \mid (f~\vec{x} \mapsto e), \Gamma \mid \epsilon & (Contexts) \\
  t &::=& \{x \mid e \} \mid x:t \to t \mid \any & (Contracts) \\
  \phi &::=& \top \mid \bot \mid x \mid \lnot \phi \mid \phi \land \phi \mid \forall x, \phi & (FOF) \\
\end{eqnarray*}

@@

$t \geq t' \iff \{ e | x \in t' \} \subset \{ e | x \in t \}$ ($t$ is more restrictive than $t'$)

case-expression can't be nested (just as lambdas) which is simpler for the FOL translation.

\section{Encoding}

\subsection{Expressions}
\subsubsection{Axioms}
\begin{itemize}
\item $\lVert x \rVert = x $
\item $\lVert f(\overline{e}) \rVert = f(\lVert \overline{e} \rVert)  $
\item $\lVert D(\overline{e}) \rVert = D(\lVert \overline{e} \rVert)  $
\item $\lVert \bad \rVert = \badphi $ (maybe we must specify that $\badphi$ is not $\top$ or $\bot$)
\item $\lVert \unr \rVert = \unrphi $ (same here)
\end{itemize}

\subsubsection{Contract satisfaction}
\begin{itemize}
\item $\llbracket \Gamma \vdash \bad \in t \rrbracket = \bot$
\item $\llbracket \Gamma \vdash \unr \in t \rrbracket = \top$
\item \begin{array}{rcll}
  \llbracket \Gamma \vdash x \in t \rrbracket &=& \top & \mbox{ if } x \mapsto t \in \Gamma \\
  && \bot & \mbox{ otherwise.} 
  \end{array}
\item \begin{array}{rcll}
  \llbracket \Gamma \vdash f\overline{(e_1,\dots,e_n)} ::: t_1 \to \dots \to t_n \to t_c \in t_c \rrbracket &=& \bigwedge_{1 \leq i \leq n}{\llbracket \Gamma \vdash e_i \in t_i \rrbracket} \\
  \llbracket \Gamma \vdash f\overline{(e_1,\dots,e_n)} ::: t \in t' \rrbracket &=& \bot \mbox{ ie type mismatch.}\\
  \end{array}
\item $\llbracket \Gamma \vdash D(\overline{e}) \in t \rrbracket = (\lVert \overline{e} \rVert \neq \badphi \leftrightarrow \lVert D(\overline{e}) \rVert \neq \badphi) \land \mbox{ check $D$ as a function application}$
\end{itemize}

Invariant:~$\llbracket \Gamma \vdash e \in t \rrbracket = \phi \iff (\phi = \top \iff \Gamma,e \mbox{ satisfies } t)$
\thispagestyle{empty}

@@

\begin{itemize}
\item $\llbracket \Gamma \vdash \mbox{let $f(\overline{x}) = $ case } e \mbox{ of } [(p_1 \to e_1),\dots,(p_n \to e_n)] \in t \rrbracket = \lVert e \rVert \neq \badphi \land (\forall \vec{z}~p_i[\vec{z}] = e \to \llbracket \Gamma, \vec{z} \mapsto \ok \vdash e_i \in t \rrbracket \lor \lVert e' \rVert = \unrphi)$ \\
  ($\vec{z}$ is a short hand for $(z_1,\dots,z_k)$ and $e'$ is the whole expression)
\item \begin{array}{rcl}
\llbracket \Gamma \vdash \mbox{ let } f(\overline{x}) = e:::t,p \rrbracket &=& \forall x_1, \dots, x_n, \{ x_1 \in t_1 \}, \dots, \{x_n \in t_n \} \\
&& \llbracket \Gamma, x_1 \mapsto t_1,\dots,x_n \mapsto t_n, f^\star \mapsto t \vdash e[f^\star/f] \in t_c \rrbracket \\
&& \land \llbracket \Gamma, f \mapsto t \vdash p \rrbracket\\
\{ x \in t \} &=& t(x) \land \forall y, t(y) \iff (\llbracket \Gamma \vdash p(y) \in \{z | z = \texttt{True} \} \rrbracket)
\end{array}
\end{itemize}

($t = \{ y | p(y) \}$ and \textt{True} refers to the Haskell boolean value (and not $\top$))
